\tool: Programmable, Efficient, and Scalable \\ Secure Two-Party Computation for Machine Learning

### 2018-05-08 ccs/eval.tex
half pass

In Table~\ref{tab:nn}, 
the column ``Model size'' is the number of parameters in the
trained model.
We observe that our performance is competitive with specialized
\minion protocols, for both the LAN and the WAN settings. Further, lines
of \tool source code required is still small. 
We note that while the \minion implementation is based on the ABY framework, it does not use the framework ``off-the-shelf'' and provide tailor-made optimizations for their applications. By contrast, \tool focuses on generic
2PC, exploiting ABY's performant implementation. Optimizations such as those found in \minion are part of our future work.
\minion also reports performance results on a bigger DNN with 7
convolution layers. In \tool, this benchmark requires partitioning and
we discuss it in \sectionref{pipeeval}.

@@

We next present \bonsai~\cite{bonsai} results on three
standard datasets: character recognition (Chars4k~\cite{campos},
accuracy 74.71\%), text recognition (USPS~\cite{hull}, accuracy
94.4\%), and object categorization (WARD\\~\cite{yang}, accuracy
95.7\%). 
The \bonsai models are already over integers and no porting from
floating-point is required.
We implement the trained classifiers in \tool for all the benchmarks
from~\cite{bonsai},
and show the representative results in Table~\ref{tab:bonsai}.
Out of all the benchmarks from~\cite{bonsai}, the dataset WARD
requires the largest model. 
The column ``depth''  shows the depth of the
tree used by \bonsai. The size of \tool 
program grows with the depth of the tree, as the straightforward \tool
implementation requires a loop for each layer of the
tree\footnote{We remark here that our current language does not support functions (which we leave for future work) and with this support, LOC would be lower and independent of depth.}.

@@

\subsection{Secure code partitioning}
\label{sec:pipeeval}
\aseem{Is it worth splitting it into microbenchamrking first, which is
  Fig. 11, and then comparison with others? Right now, Fig. 11 text
  just comes at the end, should be highlight it in a paragraph or
  something?}
The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
dataset~\cite{cifar}. 
The classifier's task is to categorize colored ($32\times 32$) images
into 10 classes. A secure evaluation of this DNN needs more memory
than what is available
on our machines. Therefore, we use partitioning and divide the
computation into seven stages. 
The first step does a convolution with 64 output channels and a ReLU
activation. The next four stages together perform a convolution that
involves multiplying a $64\times 576$ matrix with a 
$576\times 1024$ matrix. The sixth stage performs a ReLU and a
convolution. The final stage has four convolutions, five ReLUs, and a
fully connected layer. The total number of lines of \tool code for
this benchmark is 336 lines.

@@

 \caption{Partitioning results for matrix factorization. The time reported by~\cite{valeriaMatrix} for this computation is 10440 seconds.}
 \label{tab:factor} 
\end{table}
Table~\ref{tab:cifar} shows the end-to-end numbers as
well as the numbers for the sixth stage, which is the heaviest. 
The number of gates are in millions, hence the suffix `m' in the last
three columns.
%For this DNN, the specialized \minion protocol takes 544 seconds and
%communicates 9272 Mb.
As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
competitive with \minion here as well. Therefore,
we believe that with partitioning, \tool can scale to large
computations while maintaining performance
competitive with existing specialized protocols. In particular, for a
large enough DNN, \minion could run
out of memory but an appropriately partitioned \tool implementation would still
succeed. For example, consider a large DNN with $n$ layers such that
the circuit for only a single layer can  fit into memory. Then, \tool
using partitioning can evaluate this large DNN in $n$ stages. To
further illustrate the need for and utility of partitioning, we ran a
series of DNNs with different layers with and without partitioning
(where partitioning was done for every layer of the DNN). For up to
5-layer DNNs, the performance of both settings was comparable, thereby
illustrating that partitioning did not cause any overhead to
performance. For networks that had 6 or more layers, performance of
the protocol without partitioning dramatically dropped, while the
protocol with partitioning continued to scale linearly. These results
are illustrated in Figure~\ref{fig:partitioning}.

@@

\subsection{Subsequent Work}
Subsequent to our work, \aseem{TODO} et al. \aseem{cite} have
presented Gazelle, a framework for specialized DNN
protocols. Gazelle use a lattice-based packed additively homomorphic
encryption scheme (PAHE) for arithmetic computations and garbled
circuits for boolean computations. Gazelle can evaluate the CNN benchmark of
Table~\ref{tab:nn} in 0.8 seconds, as opposed to
5.1 seconds taken by \tool with the ABY backend. Such advances in
cryptographic backends are orthogonal to our contributions. In
particular, once Gazelle is available, we could add it as another
cryptographic backend to \tool. Furthermore, the authors of Gazelle remark:
``A final, very interesting and ambitious line
of work would be to build a compiler that allows us to easily
express arbitrary computations and automatically factor the
computation into PAHE and two-party primitives'' -- the exact problem
that \tool solves.

>>>>

In Table~\ref{tab:nn}, 
the column ``Model size'' is the number of parameters in the
trained model.
We observe that our performance is competitive with specialized
\minion protocols, for both the LAN and the WAN settings. Further, lines
of \tool source code required is still small. 
%We note that while the \minion implementation is based on the ABY framework, it does not use the framework ``off-the-shelf'' and provide tailor-made optimizations for their applications. %By contrast, \tool focuses on generic
%2PC, exploiting ABY's performant implementation. Optimizations such as those found in \minion are part of our future work.
\minion also reports performance results on a bigger DNN with 7
convolution layers. In \tool, this benchmark requires partitioning and
we discuss it in \sectionref{pipeeval}.

@@

We next present \bonsai~\cite{bonsai} results on three
standard datasets: character recognition (Chars4k~\cite{campos},
accuracy 74.71\%), text recognition (USPS~\cite{hull}, accuracy
94.4\%), and object categorization (WARD\\~\cite{yang}, accuracy
95.7\%). 
%The \bonsai models are already over integers and no porting from
%floating-point is required.
We implement the trained classifiers in \tool for all the benchmarks
from~\cite{bonsai},
and show the representative results in Table~\ref{tab:bonsai}.
Out of all the benchmarks from~\cite{bonsai}, the dataset WARD
requires the largest model. 
The column ``depth''  shows the depth of the
tree used by \bonsai. The size of \tool 
program grows with the depth of the tree, as the straightforward \tool
implementation requires a loop for each layer of the
tree\footnote{We remark here that our current language does not support functions (which we leave for future work) and with this support, LOC would be lower and independent of depth.}.

@@

\subsection{Secure code partitioning}
\label{sec:pipeeval}
%\aseem{Is it worth splitting it into microbenchamrking first, which is
 % Fig. 11, and then comparison with others? Right now, Fig. 11 text
 % just comes at the end, should be highlight it in a paragraph or
 % something?}
The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
dataset~\cite{cifar}. 
The classifier's task is to categorize colored ($32\times 32$) images
into 10 classes. A secure evaluation of this DNN needs more memory
than what is available
on our machines. Therefore, we use partitioning and divide the
computation into seven stages. 
The first step does a convolution with 64 output channels and a ReLU
activation. The next four stages together perform a convolution that
involves multiplying a $64\times 576$ matrix with a 
$576\times 1024$ matrix. The sixth stage performs a ReLU and a
convolution. The final stage has four convolutions, five ReLUs, and a
fully connected layer. The total number of lines of \tool code for
this benchmark is 336 lines.

@@

 \caption{Partitioning results for matrix factorization. The time reported by~\cite{valeriaMatrix} for this computation is 10440 seconds.}
 \label{tab:factor} 
\end{table}
Table~\ref{tab:cifar} shows the end-to-end numbers as
well as the numbers for the sixth stage, which is the heaviest. 
The number of gates are in millions, hence the suffix `m' in the last
three columns.
%For this DNN, the specialized \minion protocol takes 544 seconds and
%communicates 9272 Mb.
As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
competitive with \minion here as well. Therefore,
we believe that with partitioning, \tool can scale to large
computations while maintaining performance
competitive with existing specialized protocols. In particular, for a
large enough DNN, \minion could run
out of memory but an appropriately partitioned \tool implementation would still
succeed.

\paragraph{Scalability.}  To
illustrate the scalability of partitioning, we evaluate a
sequence of DNNs with and without partitioning
(where every layer of each DNN is a separate stage) in~\figureref{partitioning}.
For DNNs with up to 4 (hidden) layers, the performance is almost idential and the lines overlap, thereby
illustrating that partitioning did not cause any noticeable performance overheads.
Memory issues start showing up with the 5 (hidden) layer DNN and the non-partitioned implementation is slower.
DNNs with 6 or more layers fail to execute (terminate with a ``bus error").
However, the partitioned implementation scales well to even these large DNNs.

@@

\subsection{Subsequent Work}
Subsequent to our work, Juvekar et al.~\cite{gazelle}  have
presented {\sc Gazelle}, a specialized protocol for DNNs.
{\sc Gazelle} use a lattice-based packed additively homomorphic
encryption scheme (PAHE) for arithmetic computations and garbled
circuits for boolean computations. {\sc Gazelle} can evaluate the CNN benchmark of
Table~\ref{tab:nn} in 0.8 seconds, as opposed to
5.1 seconds taken by \tool with the ABY backend. Such advances in
cryptographic backends are orthogonal to our contributions. In
particular, once {\sc Gazelle} is available, we could add it as another
cryptographic backend to \tool. Furthermore, the authors of Gazelle remark:
``A final, very interesting and ambitious line
of work would be to build a compiler that allows us to easily
express arbitrary computations and automatically factor the
computation into PAHE and two-party primitives'' -- the exact problem
that \tool solves.

### 2018-05-08 ccs/ex.tex
nit
half pass

Using both arithmetic and boolean representations requires conversions between them.
The \tool compiler also
instruments these conversion gates accordingly. For example, in
lines~\ref{line:convacc} and~\ref{line:convb}, the compiler converts
\ls{a\_acc} and \ls{a\_b} to boolean representation, before they are
input to the comparison and multiplexer circuits. \aseem{Question: why
  is b arithmetic to begin with? Can this input be boolean and then we
  don't have to insert a conversion gate for it?}
%\vspace{-0.2in}

@@

//convert acc and b from arithmetic to boolean
$\mathtt{
share\;*y\_acc = ycirc\rightarrow PutA2YGate(a\_acc);~\label{line:convacc}
}$
$\mathtt{
share\;*y\_b = ycirc\rightarrow PutA2YGate(a\_b);~\label{line:convb}
}$

@@

\subsection{Comparison with Yao-only approach}
\aseem{We decided to put it here, but reading now, this subsection
  seems to be a bit misplaced to me. How about moving it to 6.1, the
  first subsection in the
  Evaluation? And we can add a couple of lines about it in the
  introduction evaluation paragraph (We first show that mixed
  protocols are much more efficient than Yao only by comparing with
  blah blah). Here it seems quite abrupt, we
  are talking about EzPC at a high-level, with compiler features
  etc. and suddenly details about numbers, CIFAR and what not start
  coming. Seems too soon.}
Before we provide the formalization of our compiler, we show that for performance it it critical to use a mix of arithmetic and boolean circuits.
We demonstrate the performance benefits of mixed computation over boolean-only compute by comparing with previous tools CBMC-GC~\cite{cbmcgc} and ObliVM~\cite{oblivm} that are based on Yao's garbled circuits. As we already mentioned,  we are not the first to observe that Yao-only approach does not scale well beyond examples with small numbers of multiplications \cite{aby,tasty,autoS,secureml,minionn,chameleon}, especially for large-scale multiplications commonly found in Machine Learning (ML) algorithms -- the main use case of \tool.  
Roughly, multiplying two $64$-bit integers using arithmetic sharing requires only $2$ multiplications, whereas Yao requires at least $4096$ AES operations in the online phase.
Indeed, this is one of the main drivers behind development of various specialized MPC protocols in previous works \cite{secureml,minionn,shafindss} that we compare against in \sectionref{eval}.

The largest example benchmark in CBMC-GC is multiplication of two $8\times 8$ matrices for which it generates about a million gates and takes about ten seconds to run. In contrast, in \tool we use arithmetic sharing for this compute that generates 1218 gates and runs in about 0.1 seconds. We tried running CBMC-GC for multiplying two $20\times 20$ matrices and it timed out after 5 hours. In contrast, most of our benchmarks are much bigger. For example, we run the secure computation of a DNN for CIFAR-10 in 265 seconds which requires multiplying a $64\times576$ matrix with a $576\times1024$ matrix as a sub-computation. On the other hand, ObliVM scales to larger benchmarks with its use of Yao-pipelining \cite{yao-pipe}. We ran inner-product computations of varying lengths followed by a comparison in both ObliVM and \tool (see Table~\cite{oblivm-compare} in the Appendix). 
Our experiments show that we are atleast 30x faster than ObliVM due to our use of arithmetic sharing for inner-product computation.

@@

\section{Cryptographic cost-awareness}

In this section we explain various heuristics that \tool uses to split
the secret computation in a program between arithmetic and
boolean. While an optimal split in general is an NP-hard problem
\aseem{can we cite something?}, \tool heuristics are empirically based
on micro-benchmarking, and perform well on the various secure
prediction benchmarks (Section~\ref{sec:eval}).

\paragraph{Micro-benchmarking.} The cost of splitting the compuation
between arithmetic and boolean involves the cost of operations
(addition, multiplication, etc.) and the cost of inter-conversion
gates between arithmetic and boolean. Cryptographic cost of converting
an arithmetic share to a boolean share requires computing a garbled
circuit for addition that is linear in size. Similarly, the cost of
converting from boolean to arithmetic is computing a subtraction
circuit inside garbled circuit that is again linear in size. Demmler
et al.~\cite{aby} analyze all these costs by running
microbenchmarks for basic operations and interconversions (Figures
2 and 3 in \cite{aby}). \tool heuristics are based on these results.

\paragraph{Multiplication.} For multiplication, \tool always chooses
arithmetic circuit, as the cost of boolean multiplication is much
higher than the cost of coercing the operands from boolean to
arithmetic, performing an arithmetic multiplication, and then coercing
the result back to boolean ($\sim9x$ more time $\&$ $\sim20x$ more
communication in a LAN setting). Since the performance gap is quite
large, we believe that this choice is optimal for realistic network
settings.

\paragraph{Addition.} However, this is not the case for addition. If
both the operands of addition are boolean shared, \tool performs
addition using a boolean circuit, else it uses an arithmetic
circuit. \aseem{What if one operand is in boolean and the other is in
  arithmetic?} Hence, addition may be done in arithmetic or boolean
depending on the operands.

\paragraph{Boolean operations.} Finally, \tool chooses Yao for for all
the boolean operations such as comparison, shifts, multiplexors,
etc. \aseem{Can we write a sentence here that Demmler et al. show this
  is the optimal choice?}

>>>>

Using both arithmetic and boolean representations requires conversions between them.
The \tool compiler also
instruments these conversion gates accordingly. For example, in
line~\ref{line:convacc}, the compiler converts
\ls{a\_acc} to a boolean representation, before it is
input to the comparison and multiplexer circuits.
%\vspace{-0.2in}

@@

//convert acc and b from arithmetic to boolean
$\mathtt{
share\;*y\_acc = ycirc\rightarrow PutA2YGate(a\_acc);~\label{line:convacc}
}$

@@

\subsection{Comparison with Yao-only approach}
\label{sec:oblivm}
We show that it is critical to use a mix of arithmetic and boolean circuits  for performance.
 Previous work has observed that Yao's garbled circuits do not scale to machine learning examples that require a large number of multiplications~\cite{aby,tasty,autoS,secureml,minionn,chameleon}.
Indeed, this is one of the main drivers behind the development of various specialized MPC protocols in previous works~\cite{secureml,minionn,shafindss} (we compare against them in~\sectionref{eval}) .
Here, we empirically demonstrate the performance benefits of mixed computations over boolean-only compute by comparing with state of the art Yao-based compilers CBMC-GC~\cite{cbmcgc} and ObliVM~\cite{oblivm}.

The largest benchmark in CBMC-GC is a multiplication of two $8\times 8$ matrices for which it generates about a million gates and takes about ten seconds to run. In contrast, \tool uses arithmetic sharing, generates 1218 gates, and runs in less than 0.1 seconds. When we tried multiplying two $20\times 20$ matrices with CBMC-GC, it timed out after 5 hours. Our benchmarks require much bigger computations (e.g., multiplying a $64\times576$ matrix with a $576\times1024$ matrix). Unlike CBMC-GC, ObliVM can scale to larger benchmarks (because it uses Yao-pipelining \cite{yao-pipe}). We evaluated the program in \figureref{ex-sml} with vectors of lengths varying between $10^3$ and $10^6$ using both \tool and ObliVM. 
\tool evaluates the inner-product in arithmetic and, empirically, is at least 25x faster than ObliVM.

@@

\section{Cryptographic cost-awareness}
 \label{sec:ccc}
In this section, we explain various heuristics that \tool uses to split
the computation in a program into arithmetic and
boolean parts. While finding an optimum split  is an NP-hard problem (the predicates in {\tt if} statements can have arbitrary non-linear arithmetic),
\tool uses heuristics based
on empirical micro-benchmarking that  perform well in practice (Section~\ref{sec:eval}).

The split
between arithmetic and boolean requires knowing the cost of
individual operations
(addition, multiplication, inter-conversion between arithmetic and boolean, etc.).
Demmler
et al.~\cite{aby} document these costs by running
microbenchmarks for basic operations and interconversions (Figures
2 and 3 in \cite{aby}). \tool heuristics are based on their results.

 Converting
an arithmetic share to a boolean share requires computing a garbled
circuit for addition. The size of this circuit grows  linearly in the bitwidth of the inputs. Similarly, converting from boolean to arithmetic requires computing a 
garbled circuit for subtraction, which is also linear. 
Since each conversion has roughly the same cost as a boolean addition,
 \tool performs
addition using a boolean circuit
if the operands  are boolean shared, else it uses an arithmetic
circuit.

For multiplication, \tool always chooses
an arithmetic circuit, as the cost of a boolean multiplication is much
higher than the cost of converting the operands from boolean to
arithmetic, performing an arithmetic multiplication, and then converting
the result back to boolean ($\sim9x$ more time $\&$ $\sim20x$ more
communication in a LAN setting). Theoretically, the size of a boolean multiplication circuit is quadratic in the bitwidth, which causes this performance gap.
Roughly, multiplying two $64$-bit integers using arithmetic sharing requires only $2$ multiplications, whereas Yao requires at least $4096$ AES operations in the online phase.
Since this gap is quite large, we believe that this choice is optimal for realistic network
settings. Finally, \tool chooses boolean circuits for all
the operations lacking arithmetic support in ABY, e.g., comparisons, bit-shifts, etc.

### 2018-05-08 ccs/impl.tex
half pass

We use an off-the-shelf solver
(SeaHorn~\cite{seahorn}) to bound check the array indices
($\models e < n$
in ({\sc {T-Read}}) and ({\sc{T-Write}}),
\figureref{compile}). We take the \tool source program and
translate it as an input C program to the solver. The solver takes less
than a minute on our largest benchmark to verify that all the array
accesses are in-bounds. This C program can also enable
validation of \tool generated protocols via differential testing~\cite{mckeeman,frigate}.

@@

The compilation rules of \figureref{compile} can introduce
repeated coercions from arithmetic to
boolean and vice versa.
Since \tool is aware of the cryptographic costs associated with these coercions,
it tries to minimize them using several optimizations.
For instance, if boolean shares of a variable are available in scope and the variable is involved in
an addition then \tool performs the addition in boolean rather than first coercing boolean to arithmetic
and then performing an arithmetic addition. 
%However, multiplication is always performed in arithmetic as the cost of a boolean multiplication is much higher than the cost of coercing from boolean to arithmetic, performing an arithmetic multiplication, and then coercing from arithmetic to boolean~\cite{aby}.
Other optimizations include the standard ``common subexpression elimination"
optimization~\cite{dragonbook}.
On each coercion, \tool memorizes the pair of arithmetic
share and boolean share involved in the coercion. 
\tool invalidates such pairs when the variables corresponding to the
shares are overwritten by assignments. 
In subsequent coercions, \tool  reuses valid pairs (if available)
instead of inserting code to recompute them afresh.

All these optimizations are standard compiler
optimizations~\cite{dragonbook}, and we rely on their correctness to
maintain the correctness and security for the optimized programs.

>>>>

We use an off-the-shelf solver
(SeaHorn~\cite{seahorn}) to check that the array indices
are within bounds ($\models e < n$
in ({\sc {T-Read}}) and ({\sc{T-Write}}),
\figureref{compile}). We take the \tool source program and
translate it as an input C program to the solver. The solver takes less
than a minute on our largest benchmark to verify that all the array
accesses are in-bounds. This C program also enables
validation of \tool generated protocols via differential testing~\cite{mckeeman,frigate}.

@@

The compilation rules of \figureref{compile} can introduce
repeated coercions from arithmetic to
boolean and vice versa.
Since \tool is aware of the cryptographic costs associated with these coercions,
it tries to minimize them using several optimizations, e.g., by
%For instance, if boolean shares of a variable are available in scope and the %variable is involved in
%an addition then \tool performs the addition in boolean rather than first %coercing boolean to arithmetic
%and then performing an arithmetic addition. 
%However, multiplication is always performed in arithmetic as the cost of a %boolean multiplication is much higher than the cost of coercing from boolean %to arithmetic, performing an arithmetic multiplication, and then coercing from %arithmetic to boolean~\cite{aby}.
%Other optimizations include
the standard ``common subexpression elimination"
optimization~\cite{dragonbook}.
On each coercion, \tool memorizes the pair of arithmetic
share and boolean share involved in the coercion. 
\tool invalidates such pairs when the variables corresponding to the
shares are overwritten by assignments. 
In subsequent coercions, \tool  reuses valid pairs (if available)
instead of inserting code to recompute them afresh.
These optimizations are standard compiler
optimizations~\cite{dragonbook}, and we rely on their correctness to
maintain the correctness and security of the optimized programs.

### 2018-05-08 ccs/ld.tex
pass over paper. added partitioning figure.
half pass

In this section we prove correctness and security of our \tool compiler.
The readers who are interested in using the compiler as a black box can
move directly to  Section~\ref{sec:pipe}.
%\dg{Put a warning that this section is technical and not prerequisite for understanding later section.}
%We formalize our source, intermediate and circuit languages. 
%An example of a program in our source language is provided in \figureref{ex-sml}.
%Our source runtime semantics is a model of the ideal, trusted third-party semantics, and
%generates observations corresponding to the values revealed to the
%parties.
We first formalize our source language (an example program being
\figureref{ex-sml}), and its runtime semantics.
This semantics describes the ``trusted third party"
execution semantics of the source programs and
generates observations corresponding to the values revealed to the
parties.
%
We then present the compilation rules that type check a program in the
source language and generate a program in the intermediate
language (an example program being \figureref{ex-aby}).
%
Next, we present the runtime semantics of our intermediate language
that evaluates to a circuit by ``evaluating away'' the public parts
and the arrays. Crucially, this step does not have access to the secret
inputs; those are processed by our distributed circuit semantics that
model the \mpc back-end.
%
Evaluation in this distributed setting involves the parties running an
interactive protocol. This step, like  the source semantics, emits
observations corresponding to the values revealed to the parties.
%Finally, we formalize the distributed circuit semantics (of the cryptographic back-end) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observationscorresponding to the values revealed to the parties.

@@

\paragraph{Source language.} Our source language
 is a simple imperative language shown in \figureref{srclang}. Types
$\psi$ consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports
 multi-dimensional arrays as well. Expressions $e$ in the language include the
 integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} >
e_{2}$ (we support several other
operators in the implementation, detailed in \sectionref{impl}),
conditionals $\cond{e}{e_{1}}{e_{2}}$,
array literals
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations, assignments,
\kw{for} loops, array writes, \kw{if}
statements, and sequence of statements. The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties\footnote{Our language also has statements $\kw{out}_1\:e$ (resp., $\kw{out}_2\:e$) to reveal value of $e$ to only the first (resp., second) party. We omit these for brevity.}. The \kw{while} statement is an internal syntax that is not
exposed to the programmer. \\

>>>>

In this section we prove correctness and security of our \tool compiler.
The readers who are interested in using the compiler as a black box can
move directly to  Section~\ref{sec:pipe} without loss of continuity.
%\dg{Put a warning that this section is technical and not prerequisite for understanding later section.}
%We formalize our source, intermediate and circuit languages. 
%An example of a program in our source language is provided in \figureref{ex-sml}.
%Our source runtime semantics is a model of the ideal, trusted third-party semantics, and
%generates observations corresponding to the values revealed to the
%parties.
We first formalize our source language (an example program being
\figureref{ex-sml}), and its runtime semantics.
This semantics describes the ``trusted third party"
execution semantics of the source programs and
generates observations corresponding to the values revealed to the
parties.
%
We then present the compilation rules that type check a program in the
source language and generate a program in the intermediate
language (an example program being \figureref{ex-aby}).
%
Next, we present the runtime semantics of our intermediate language
that evaluates to a circuit by ``evaluating away'' the public parts
and the arrays. Crucially, this step does not have access to the secret
inputs; those are processed by our distributed circuit semantics that
model the \mpc back-end.
%
Evaluation in this distributed setting involves the parties running an
interactive protocol. This step, like  the source semantics, emits
observations corresponding to the values revealed to the parties.
%Finally, we formalize the distributed circuit semantics (of the cryptographic back-end) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observationscorresponding to the values revealed to the parties.

@@

\paragraph{Source language.} Our language
 is a simple imperative language shown in \figureref{srclang}. Types
$\psi$ consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports
 multi-dimensional arrays as well. Expressions $e$ in the language include the
 integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} >
e_{2}$ (we support several other
operators in the implementation, detailed in \sectionref{impl}),
conditionals $\cond{e}{e_{1}}{e_{2}}$,
array literals
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations, assignments,
\kw{for} loops, array writes, \kw{if}
statements, and sequence of statements. The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties\footnote{Our language also has statements $\kw{out}_1\:e$ (resp., $\kw{out}_2\:e$) to reveal value of $e$ to only the first (resp., second) party. We omit these for brevity.}. The \kw{while} statement is an internal syntax that is not
exposed to the programmer. \\
